### 2018-07-02 paper/7-related.tex
S7 adjust fabricate
Work on S7

\begin{itemize}
\item \Ninja~\cite{ninja} combines the dependency strategy of \Make with the
validating traces of \Shake~--~our associated implementation provides such a
combination. \Ninja~is also capable of modelling rules that produce multiple
results, a limited form of polymorphism \S\ref{sec-polymorphism}.

@@

\item \CloudBuild~\cite{esfahani2016cloudbuild} uses constructive traces with a topological sort.

\item \Buck~\cite{buck} uses deep constructive traces with a topological sort,
meaning that it disallows non-determinism~\S\ref{sec-non-determinism}.

\item \Nix~\cite{dolstra2004nix} has coarse-grained dependencies, with precise
hashing of dependencies and downloading of precomputed build products. When
combined with \cmd{import-from-derivation}, \Nix provides monadic dependencies.
\Nix is not primarily intended as a build system, and the coarse grained nature
(packages, not individual files) makes it targeted to a different purpose.
% John Ericson suggested in a blog comment that nix may be somewhat monadic, see:
% \url{https://blogs.ncl.ac.uk/andreymokhov/cloud-and-dynamic-builds/\#comment-1849}.

@@

The one build system we are aware of that cannot be modelled in our framework is
\Fabricate~\cite{fabricate}. In \Fabricate a build system is a script which is
run in-order, in the spirit of\footnote{\Fabricate requires scripts to be
written in Python, but those details are not fundamental to what makes
\Fabricate special.}:

@@

\noindent
To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed.

Taking our abstraction, it is possible to encode \Fabricate assuming that
commands like \cmd{gcc -c util.c} are keys, there is a linear dependency between
each successive node, and that the OS-level tracing can be lifted back as a
monadic \hs{Task} function\footnote{In fact, \Shake has an execution mode that
can model \Fabricate{}-like build systems~--~see \hs{Development.Shake.Forward}
in the \Shake library.}. However, in our pure model the mapping is not perfect
as \cmd{gcc} writes to arbitrary files whose locations are not known in advance.

@@

While not typically considered build systems, self-adjusting computation is a
well studied area, and in particular the contrast between different formulations
has been thoroughly investigated~\cite{acar2007selfadjusting}.

Self-adjusting computations can automatically adjust to an external change
to their inputs. A classic example is a self-adjusting sorting algorithm, which
can efficiently (in $O(\log{n})$ time where $n$ is the length of the input)
recalculate the result given an incremental change of the input. While very
close to build systems in spirit, self-adjusting computations are mostly used
for in-memory computation and rely on the ability to dynamically allocate new
keys in the store for sharing intermediate computations~--~an intriguing feature
rarely seen in build systems (\Shake's oracles~\S\ref{sec-polymorphism} can be
used to model this feature to a limited degree).

@@

\begin{minted}[xleftmargin=10pt]{haskell}
type Task c k v = @\std{forall}@ f. c f => (k -> f v) -> k -> Maybe (f v)
\end{minted}

>>>>

\begin{itemize}
\item \Ninja~\cite{ninja} combines the \hs{topological} scheduler of \Make with
the verifying traces of \Shake~--~our associated implementation provides such a
combination. \Ninja~is also capable of modelling build rules that produce
multiple results, a limited form of polymorphism \S\ref{sec-polymorphism}.

@@

\item \Nix~\cite{dolstra2004nix} has coarse-grained dependencies, with precise
hashing of dependencies and downloading of precomputed build products. We
provided a model of \Nix in \S\ref{sec-implementation-cloud}, although it is
worth noting that \Nix is not primarily intended as a build system, and the
coarse grained nature (packages, not individual files) makes it targeted to a
different purpose.

@@

The one build system we are aware of that cannot be modelled in our framework is
\Fabricate by \citet{fabricate}. In \Fabricate a build system is a script that is
run in-order, in the spirit of:
% I'm not sure the footnote adds much, let's save space?
% \footnote{\Fabricate requires scripts to be
% written in Python, but those details are not fundamental to what makes
% \Fabricate special.}

@@

To achieve minimality, each separate command is traced at the OS-level, allowing
\Fabricate to record a trace entry stating that \cmd{gcc -c util.c} reads from
\cmd{util.c}. In future runs \Fabricate runs the script from start to finish,
skipping any commands where no inputs have changed.

Taking our abstraction, it is possible to encode \Fabricate assuming that
commands like \cmd{gcc -c util.c} are keys, there is a linear dependency between
each successive key, and that the OS-level tracing can be lifted back as a
monadic \hs{Task} function\footnote{\Shake provides support for
\Fabricate{}-like build systems~--~see \cmd{Development.Shake.Forward}
in the \Shake library.}. However, in our pure model the mapping is not perfect
as \cmd{gcc} writes to arbitrary files whose locations are not known in advance.

@@

While not typically considered build systems, self-adjusting computation is a
well studied area, and in particular the contrast between different formulations
has been thoroughly investigated, e.g. see~\cite{acar2007selfadjusting}.
Self-adjusting computations can automatically adjust to an external change
to their inputs. A classic example is a self-adjusting sorting algorithm, which
can efficiently (in $O(\log{n})$ time where $n$ is the length of the input)
recalculate the result given an incremental change of the input. While very
close to build systems in spirit, self-adjusting computations are mostly used
for in-memory computation and rely on the ability to dynamically allocate new
keys in the store for sharing intermediate computations~--~an intriguing feature
rarely seen in build systems (\Shake's oracles~\S\ref{sec-polymorphism} can be
used to model this feature to a limited degree).

@@

\begin{minted}[xleftmargin=10pt]{haskell}
newtype Task c k v = Task { run :: @\std{forall}@ f. c f => (k -> f v) -> f v }
\end{minted}
